---
title: "Getting Started with Binomial"
author: "Frankie Lin"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with Binomial}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Binomial

The package `"binomial"` is a basic implementation for calculating various values such as mean, mode, probability etc, and a way to visualize the probability of successes within the discrete binomial framework. 

## Loading Package
```{r}
library(binomial)
```


## Finding the combinations and probability of binomial distribution

The first step is to understand the ability to calulates the combinations and the probability of a given binomial event. 

```{r}
# bin_choose function
bin_choose(n = 10, k = 5)

#bin_probability function
bin_probability(success = 5, trials = 10, prob = 0.5)
```

## Binomial distribution

Additionally to just calculating a single event, the package offers a variety of methods of pulling the modifying distribution of a given number of trials. In addition to that, our package allows you to easily plot these distributions. 

```{r}
# Binomial Distribution
bin_distribution(10, 0.5)

# Plotting
plot(bin_distribution(10, 0.5))
```

Additionally to standard distributions, one can also look at cumulative distributions:
```{r}
# Culumlative Distribution
bin_cumulative(10, 0.5)

# Plotting
plot(bin_cumulative(10, 0.5))
```

Moreso, we can print the summary statistics for particuliar binomial probabilitie by defining a binomial variable:
```{r}
# defining a binomial variable
variable <- bin_variable(10, .5)

# Priting general summary
variable

# Printing longform summary
var.summary <- summary(variable)
var.summary
```

## Misc. Calculations

With this package there are quite a few functions that be used to calculate different values displayed in the summary above. 
```{r}
bin_mean(10, 0.5)
bin_variance(10, 0.5)
bin_mode(10, 0.5)
bin_kurtosis(10, 0.5)
bin_skewness(10, 0.5)
```

